systemfit: A Package to Estimate\\ Simultaneous Equation Systems in \proglang{R}

### 2005-09-26 paper/trunk/systemfit.tex
some minor changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@573 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\section{CVS Revision Log}
\begin{verbatim}
$Log$
Revision 1.11  2005/09/01 12:27:55  henningsena
revised structure of the document (sections, subsections, subsubsections) due to removal of nlsystemfit; moved each section into a separate file.

>>>>

\section{CVS Revision Log}
\begin{verbatim}
$Log$
Revision 1.12  2005/09/26 11:24:05  henningsena
some minor changes

Revision 1.11  2005/09/01 12:27:55  henningsena
revised structure of the document (sections, subsections, subsubsections) due to removal of nlsystemfit; moved each section into a separate file.

### 2005-09-26 paper/trunk/systemfit_estimation.tex
some minor changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@573 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ordinary Least Squares estimation}

@@

If no cross-equation restrictions are imposed, the simultaneous 
OLS estimation of the system leads to the same parameter estimates 
as an equation-wise OLS estimation. The variance-covariance matrix 
of the parameters from an equation-wise OLS estimation is equal to 
the variance-covariance matrix obtained by equation 
(\ref{eq:OLSvcm2}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Weighted least squares}

@@

If no cross-equation restrictions are imposed, 
the parameter estimates and the variance-covariance matrix 
of the parameters are equal to the OLS estimates%
\footnote{The variance-covariance matrices of the 
estimated parameters might differ, 
if the WLS estimation is based on a different formula 
to calculate the (diagonal) variance-covariance 
matrix of the residuals (see section \ref{sec:residcov}).}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Seemingly unrelated regression}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Two-stage least squares estimation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Weighted two-stage least squares estimation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Three-stage least squares estimation}

@@

The econometrics software EViews uses following approach:
\begin{equation}
   \widehat{\beta} = \widehat{\beta}_{2SLS} + 
   \left( \widehat{X}' \Omega^{-1} \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}_{2SLS} \right)
\end{equation}
with $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
(equation \ref{eq:beta2sls}).
EViews uses the standard 3SLS formula (equation (\ref{eq:cov3sls})) to 
calculate the variance-covariance matrix of the 3SLS estimator.

If the same instrumental variables are used in all equations 
($H_1 = H_2 = \ldots = H_G$), 
all the above mentioned approaches lead to identical parameter estimates.
However, if this is not the case, the results depend on the 
equation used \citep{schmidt90}. 
The only reason to use different instruments in the different equations 
is that the instruments of one equation are correlated with the
disturbance terms in another equation. Otherwise, one would simply use
every instrument in every equation \citep{schmidt90}. 
In this case, only the GMM-3SLS (equation (\ref{eq:beta3slsGMM}))
and the 3SLS estimator developed by \cite{schmidt90} 
(equation (\ref{eq:beta3slsSchmidt})) are consistent.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ordinary least squares (OLS)}

@@

If no cross-equation parameter restrictions are imposed, the simultaneous 
OLS estimation of the system leads to the same parameter estimates 
as an equation-wise OLS estimation. The variance-covariance matrix 
of the parameters from an equation-wise OLS estimation is equal to 
the variance-covariance matrix obtained by equation 
(\ref{eq:OLSvcm2}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Weighted least squares (WLS)}

@@

If no cross-equation parameter restrictions are imposed, 
the parameter estimates are equal to the OLS estimates.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Seemingly unrelated regression (SUR)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Two-stage least squares (2SLS)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Weighted two-stage least squares (W2SLS)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Three-stage least squares (3SLS)}

@@

The econometrics software EViews uses following approach:
\begin{equation}
   \widehat{\beta} = \widehat{\beta}_{2SLS} + 
   \left( \widehat{X}' \Omega^{-1} \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}_{2SLS} \right)
\end{equation}
with $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
(equation \ref{eq:beta2sls}).
EViews uses the standard 3SLS formula (\ref{eq:cov3sls}) to
calculate the variance-covariance matrix of the 3SLS estimator.

If the same instrumental variables are used in all equations 
($H_1 = H_2 = \ldots = H_G$), 
all the above mentioned approaches lead to identical parameter estimates.
However, if this is not the case, the results depend on the 
equation used \citep{schmidt90}. 
The only reason to use different instruments in the different equations 
is that the instruments of one equation are correlated with the
disturbance terms in another equation. Otherwise, one would simply use
every instrument in every equation \citep{schmidt90}. 
In this case, only the GMM-3SLS (\ref{eq:beta3slsGMM})
and the 3SLS estimator developed by \cite{schmidt90} 
(\ref{eq:beta3slsSchmidt}) are consistent.

### 2005-09-26 paper/trunk/systemfit_restrictions.tex
some minor changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@573 591db2bd-e0a0-4d22-8c19-2192dbb7da31

There are two ways to impose linear parameter restrictions.
First, a matrix $T$ can be provided with
\begin{equation}
   \beta = T \cdot \beta^* \label{eq:T-restr} 
\end{equation}
where $\beta^*$ is a vector of linear independent coefficients.

@@

The first way is less flexible than this latter one%
\footnote{e.g. restrictions like $\beta_1 + \beta_2 = 4$ cannot be imposed
by the first method}, 
but the first way is preferable if equality constraints for coefficients
across many equations of the system are imposed. 
Of course, these restrictions can be also imposed using
the latter method.
However, while the latter method increases the dimension of the 
matrices to be inverted during estimation, the first reduces it. 
Thus, in some cases the latter way leads to estimation problems
(e.g. (near) singularity of the matrices to be inverted),
while the first doesn't.

@@

The econometrics software EViews calculates the restricted 3SLS estimator by:
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta^0}_{2SLS} \right) 
      \\ q 
   \end{array} \right]
\end{equation}
where $\widehat{\beta^0}_{2SLS}$ is the restricted 2SLS estimator calculated
by equation (\ref{eq:beta2SLSr}). 
To calculate the variance-covariance matrix 
EViews uses the standard formula of the restricted 3SLS estimator
(equation (\ref{eq:cov3slsr})).

>>>>

There are two ways to impose linear parameter restrictions.
First, a matrix $T$ can be provided with
\begin{equation}
   \beta = T \cdot \beta^* \label{eq:T-restr} 
\end{equation}
where $\beta^*$ is a vector of restricted (linear independent) coefficients,
and $T$ is a matrix with the number of rows equal to the number of
unrestricted coefficients ($\beta$) and
the number of columns equal to the number of restricted coefficients
($\beta^*$).
$T$ can be used to map each unrestricted coefficient to one or more
restricted coefficients.

@@

The first way is less flexible than this latter one%
\footnote{e.g.\ restrictions like $\beta_1 + \beta_2 = 4$ cannot be imposed
by the first method.}, 
but the first way is preferable if equality constraints for coefficients
across many equations of the system are imposed. 
Of course, these restrictions can be also imposed using
the latter method.
However, while the latter method increases the dimension of the 
matrices to be inverted during estimation, the first reduces it. 
Thus, in some cases the latter way leads to estimation problems
(e.g. (near) singularity of the matrices to be inverted),
while the first doesn't.

@@

The econometrics software EViews calculates the restricted 3SLS estimator by:
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta^0}_{2SLS} \right) 
      \\ q 
   \end{array} \right]
\end{equation}
where $\widehat{\beta^0}_{2SLS}$ is the restricted 2SLS estimator calculated
by equation (\ref{eq:beta2SLSr}). 
To calculate the variance-covariance matrix 
EViews uses the standard formula of the restricted 3SLS estimator
(\ref{eq:cov3slsr}).
